
## Replicated and consensus peaks

### Replicated peaks {.tabset .tabset-pills}

For each group (i.e. replicates from the same sample group and target as listed in section **Sample sheet**), peaks from all replicates are pooled and merged into a set of non-overlapping peaks.\
If multiple replicates are available, only **replicated peaks** contributed by at least two replicates (or more replicates if a specific cutoff is applied) are kept for further analysis.\
If only one replicate is available or only one replicate produces peaks, its peaks is used for further analysis unless excluded specifically otherwise.\
Peaks that are produced from this step, either replicated peaks or peaks from the only one replicate, are referred to as **group peaks**, i.e. one set of peaks for each *group* as listed in the sample sheet, and are used to generate **consensus peaks** (see below). \

$$\\[0.1in]$$


```{r rep_plots, results='asis', fig.dim=dim.rep}
qc <- 'count_reps'
if (qc %in% names(figs)){
  plist <- figs[[qc]]
  for (tgt in names(plist)){
    cat("\n\n")
    cat("####", tgt, "\n\n")
    print(plist[[tgt]])
    cat("\n\n")
    
  }
  
}

```

#### Table
```{r}
if ('nreps' %in% names(dat)){
  npeaks <- dat$npeaks %>% 
    mutate(
      group = paste(sample_group, target, sep = '_'),
      peak_type = paste(toIgG, filtered, sep = '_'),
      sample_replicate = paste0('replicate', sample_replicate)
      ) %>% 
    dplyr::select(id, group, caller, peak_type, sample_replicate, npeak)
  
  ## get statistics of the final set of original peaks
  npeaks_final <- bind_rows(lapply(
    split(npeaks, paste(npeaks$id, npeaks$caller)),
    function(x){
      if(sum(grepl('_filtered', x$peak_type))){
        x %>% filter(grepl('_filtered', x$peak_type))
      }else{
        x %>% filter(grepl('Target_only_unfiltered', x$peak_type))
      }
    }
  )) %>% 
    dplyr::select(-id) %>% 
    tidyr::pivot_wider(names_from = sample_replicate, values_from = npeak) 
  
  nreps <- dat$nreps%>% 
    dplyr::select(group, caller, peak.count, ids, output_file) %>% 
    dplyr::rename("replicated.peaks" = "peak.count") %>% 
    left_join(
      npeaks_final, by = c('group', 'caller')
    ) %>% 
    dplyr::relocate(ids, peak_type, output_file, .after=last_col())
  
  
  nreps  %>% 
    DT::datatable(
      class = 'cell-border stripe', 
      rownames = T,  
      filter = 'top', 
      editable = F, 
      extensions = 'Buttons', 
      options = list(
        pageLength = 10, 
        dom = 'Bfrtip',
        buttons = c('csv')
      ))
}
```

### {.unnumbered .unlisted}

$$\\[0.2in]$$

### Consensus peaks by target {.tabset .tabset-pills}

For each target, **group peaks** from all sample groups, e.g. cell line, conditions, are pooled, and overlapping peaks are merged into a set of **consensus peaks**.\

$$\\[0.1in]$$

#### Peak count

```{r }
if ('nconps' %in% names(dat)){
data.table::rbindlist(lapply(
  split(dat$nconps, dat$nconps$target),
  function(x){
    data.frame(
      target = x$target[1],
      MACS2_narrow_peaks = ifelse('MACS2narrow' %in% x$caller, x$nconp[x$caller == 'MACS2narrow'], NA),
      MACS2_broad_peaks = ifelse('MACS2broad' %in% x$caller, x$nconp[x$caller == 'MACS2broad'], NA),
      SEACR_peaks = ifelse('SEACR' %in% x$caller, x$nconp[x$caller == 'SEACR'], NA)
    )
  }
)) %>% 
  kable %>%
  kable_styling("striped", full_width = F) %>%
  scroll_box(width = "700px", height = "200px")
}
```

#### Genomic distribution

```{r fig.dim=dim.ann }
qc <- 'conp_ann'

if (qc %in% names(figs)){
  plist <- list()
  keep.targets <- unique(gsub('.macs2.*|.seacr.*', '', names(figs[[qc]])))
  for (tgt in keep.targets){
    conps <- paste(tgt, c('macs2_narrow_peaks', 'macs2_broad_peaks', 'seacr_peaks'), sep = '.')
    for (id in conps){
      if(id %in% names(figs[[qc]])){
        plist <- c(plist, list(figs[[qc]][[id]]))
      }else{
        plist <- c(plist, list(plot_spacer()))
      }
    }
    plist <- c(plist, rep(list(plot_spacer()),3))
  }
  p <- wrap_plots(plist, ncol = 3, heights = rep(c(1,0.05), length(keep.targets)))
    
  print(p)
}

```

#### Compare sample groups

Plot distribution of consensus peaks in each sample group.\
*shared* = peaks contributed by at least 2 sample groups.\
*unique* = unqiue to the sample group
.\

```{r fig.dim=dim.conp }
qc <- 'rep2conp'
if (qc %in% names(figs)){
  plist <- list()
  keep.targets <- unique(gsub('.macs2.*|.seacr.*', '', names(figs[[qc]])))
  for (tgt in keep.targets){
    conps <- paste(tgt, c('macs2_narrow_peaks', 'macs2_broad_peaks', 'seacr_peaks'), 'bed', sep = '.')
    for (id in conps){
      if(id %in% names(figs[[qc]])){
        plist <- c(plist, list(figs[[qc]][[id]]))
      }else{
        plist <- c(plist, list(plot_spacer()))
      }
    }
    plist <- c(plist, rep(list(plot_spacer()),3))
  }
  
  p <- wrap_plots(c(plist, list(guide_area())), ncol = 3, guides = 'collect')+
    plot_layout(heights = c(rep(c(1,0.2), length(keep.targets)), 0.2))
  print(p)
  
}


```

#### MACS2 vs SEACR

Compare consensus peaks generated by MACS2 and SEACR.\

```{r fig.dim=dim.venn }
if ('seacr2macs' %in% names(figs)){
  p <- figs$seacr2macs[!sapply(figs$seacr2macs, is.null)]
  wrap_plots(p, ncol = ceiling(sqrt(ntargets)))
}
```






## {.unnumbered .unlisted}

$$\\[0.5in]$$


