### Replicated peaks {.tabset .tabset-pills}

For each sample group, peaks from all replicates were pooled and merged into non-overlapping peaks.
If multiple replicates are available, only **replicated peaks** (contributed by at least two replicates) were kept for further analysis.\

$$\\[0.1in]$$


```{r}
if ('nreps' %in% names(dat)){
  npeaks <- dat$npeaks %>% 
    mutate(
      group = paste(sample_group, target, sep = '_'),
      peak_type = paste(toIgG, filtered, sep = '_'),
      sample_replicate = paste0('replicate', sample_replicate)
      ) %>% 
    dplyr::select(id, group, caller, peak_type, sample_replicate, npeak)
  npeaks_final <- bind_rows(lapply(
    split(npeaks, paste(npeaks$id, npeaks$caller)),
    function(x){
      if(sum(grepl('_filtered', x$peak_type))){
        x %>% filter(grepl('_filtered', x$peak_type))
      }else{
        x %>% filter(grepl('Target_only_unfiltered', x$peak_type))
      }
    }
  )) %>% 
    dplyr::select(-id) %>% 
    tidyr::pivot_wider(names_from = sample_replicate, values_from = npeak) 
  
  nreps <- dat$nreps%>% 
    dplyr::select(group, caller, peak.count, ids) %>% 
    dplyr::rename("replicated.peaks" = "peak.count") %>% 
    left_join(
      npeaks_final, by = c('group', 'caller')
    ) %>% 
    dplyr::relocate(ids, peak_type, .after=last_col())
  
  
  nreps  %>% 
  DT::datatable(
  class = 'cell-border stripe', 
  rownames = T,  
  filter = 'top', 
  editable = F, 
  extensions = 'Buttons', 
  options = list(
    pageLength = 10, 
    dom = 'Bfrtip',
    buttons = c('csv')
  ))
}
```


### Consensus peaks {.tabset .tabset-pills}

For each target, peaks from all conditions were pooled, and overlapping peaks were merged into a set of consensus peaks across conditions.\

$$\\[0.1in]$$

#### Peak count

```{r conp_count}
if ('nconps' %in% names(dat)){
data.table::rbindlist(lapply(
  split(dat$nconps, dat$nconps$target),
  function(x){
    data.frame(
      target = x$target[1],
      MACS2_narrow_peaks = ifelse('MACS2narrow' %in% x$caller, x$nconp[x$caller == 'MACS2narrow'], NA),
      MACS2_broad_peaks = ifelse('MACS2broad' %in% x$caller, x$nconp[x$caller == 'MACS2broad'], NA),
      SEACR_peaks = ifelse('SEACR' %in% x$caller, x$nconp[x$caller == 'SEACR'], NA)
    )
  }
)) %>% 
  kable %>%
  kable_styling("striped", full_width = F) %>%
  scroll_box(width = "700px", height = "200px")
}
```

#### Genomic distribution

```{r fig.width=9, fig.height=fig.conp.height}
qc <- 'conp_ann'

if (qc %in% names(figs)){
  p <- wrap_plots(lapply(
    paste(rep(targets,each=3), c('macs2_narrow_peaks', 'macs2_broad_peaks', 'seacr_peaks'), sep = '.'),
    function(id){
      if(id %in% names(figs[[qc]])){
        figs[[qc]][[id]]
      }else{
        plot_spacer()
      }
    }), ncol = 3)
  
  
  print(p)
}

```

#### Compare sample groups

Plot distribution of consensus peaks in each sample group.\
*shared* = peaks contributed by at least 2 sample groups.\
*unique* = unqiue to the sample group
.\

```{r fig.width=9, fig.height=fig.conp.height2}
if ('rep2conp' %in% names(figs)){
  wrap_plots(figs$rep2conp, ncol = 3)
}


```

#### MACS2 vs SEACR

Compare consensus peaks generated by MACS2 and SEACR.\

```{r fig.width=fig.conp.width}
if ('seacr2macs' %in% names(figs)){
  p <- figs$seacr2macs[!sapply(figs$seacr2macs, is.null)]
  wrap_plots(p, ncol = ceiling(sqrt(ngroups)))
}
```




