
## QC reads {.tabset .tabset-pills}

### Plots {.tabset .tabset-pills}
**A. Sequenced reads** - number of paired reads after adapter trimming and passed to alignment.

**B. Reads aligned to target genome** - number of reads aligned to the target genome.

**C. Alignment rate to target genome** - fraction of reads aligned to the target genome.

**D. Reads aligned to E. coli genome** - number of reads aligned to E. coli genome.

**E. Duplication rate** - fraction of duplicated reads.

**F. Estimated library size** - estimated number of fragments in a library given the duplication rate.

**G. Fragment length** - fragment length estimated based on paired-end sequencing.  

```{r read_plots, results='asis', fig.width=fig.read.width, fig.height=fig.read.height}
qc_read <- c("seq_depth", "aligned_reads", "aligned_pct", "aligned_reads_spikein", "dup_rate", "est_lib_size", "frag_lens_dens")
if (length(qc_read) == 0){
  cat('No plots were found!')
  quit()
}

## targeted proteins
for (tgt in c(targets, 'IgG')){
  p1 <- wrap_plots(
    c(lapply(
      qc_read, 
      function(i){
        if (i %in% names(figs)){
          figs[[i]][[tgt]]
        }else{
          ggplot(data.frame(x=0,y=0))+
          geom_blank()+
            theme_minimal()
        }
        
      }), list(guide_area())), 
    ncol = 3, guides = 'collect')+
    plot_annotation(tag_levels = 'A')+
    plot_layout(design = c(
      area(1,1,4,9), area(1,11,4,19), area(1,21,4,29), 
      area(6,1,9,9), area(6,11,9,19), area(6,21,9,29), 
      area(11,1,19,9), area(11,11,19,19)
    ))
  
  cat("\n\n")
  cat("####", tgt, "\n\n")
  print(p1)
  cat("\n\n")
  
}


```


### Metrics
```{r}
cols2format <- intersect(c('bt2_overall_alignment_rate_target', 'bt2_overall_alignment_rate_spikein', 'dedup_percent_duplication'), colnames(meta))
meta %>% 
  mutate(
    bt2_total_aligned_Ecoli = bt2_total_aligned_spikein,
    Ecoli_alignment_rate = bt2_total_aligned_spikein/bt2_total_reads_spikein,
  ) %>% 
  arrange(sample_group, target, sample_replicate) %>% 
  relocate(id, target, sample_group, sample_replicate) %>% 
  unique.data.frame() %>% 
  DT::datatable(
  class = 'cell-border stripe', 
  rownames = T,  
  filter = 'top', 
  editable = F, 
  extensions = 'Buttons', 
  options = list(
    pageLength = 6, 
    dom = 'Bfrtip',
    buttons = c('csv')
  )) %>% 
  DT::formatRound(columns=cols2format, digits=4)

```

##  {.unlisted .unnumbered}

$$\\[0.5in]$$

